---
title: "ggplot2_solutions_chapter9"
author: "Nade Kang"
date: "July 12, 2018"
output: html_document
---

# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 9 Data Analysis

## *Exercise 9.3.3*
### Load Packages
```{r setup, results='hide'}
library(tidyverse)
library(ggthemes)
```

**_Question 1_** How can you translate each of the initial example datasets into the other
form?

I do not have the ec2 dataset. Instead, I'll use the data in tidyverse:
```{r e.9.3.3.1}
table4b
table4b_trans <- table4b %>%
  gather(key = "year", value = "population", `1999`, `2000`)
table4b_trans

table4b_trans %>% 
  spread(key = "year", value = "population")
```

**_Question 2_** How can you convert back and forth between the economics and
economics long datasets built into ggplot2?

Take a look at the economics and economics_long dataset, it is easy to discover that
the latter one is simply the gathered economics dataset.
However, I do not know how the economic_long dataset created value01 column.

```{r e.9.3.3.2}
economics
economics_long

economics_longe <- economics %>% 
  gather(pce, pop, psavert, uempmed, unemploy, key = "variable", value = "value")
economics_longe
```

**_Question 3_** Install the EDAWR package from https://github.com/rstudio/EDAWR.
Tidy the storms, population and tb datasets.

I installed the package using a R script. You have to do it your own to install it by
typing install.packages("EDAWR")

I got a warning:
Warning in install.packages :
  package ‘EDAWR’ is not available (for R version 3.4.4)
So cannot complete this question here.

## *Exercise 9.4.1*
**_Question 1_** Install the EDAWR package from https://github.com/rstudio/EDAWR.
Tidy the who dataset.

Cannot complete this one.

**_Question 2_** Work through the demos included in the tidyr package (demo(package =
"tidyr"))
```{r, eval=FALSE}
(demo(package = "tidyr"))
```

```{r e.9.4.1.2_1}
smiths <- tidyr::smiths
smiths

smiths_gather <- smiths %>% 
  gather(age, weight, height, key = "characters", value = "value")
smiths_gather
```

We can spread the population dataset:

```{r e.9.4.1.2_2}
population <- tidyr::population
head(population)

pop_spread <- population %>% 
  spread(key = "year", value = "population")
pop_spread
```

