---
title: "ggplot2_solutions_chapter7"
author: "Nade Kang"
date: "July 11, 2018"
output: html_document
---

# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 7 Positioning

## *Exercise 7.2.7*
### Load Packages
```{r setup, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

**_Question 1_** Diamonds: display the distribution of price conditional on cut and carat.
Try facetting by cut and grouping by carat. Try facetting by carat and
grouping by cut. Which do you prefer?

First we create two bases:
```{r e.7.2.7.1_base}
base_cut <- ggplot(diamonds)
base_cut
base_cut + geom_histogram(aes(x = price, fill = cut), position = "dodge")

base_carat <- ggplot(diamonds, aes(carat, price))
base_carat +geom_point()
```

Facetting cut and group by carat
```{r e.7.2.7.1_plot1}
base_carat + geom_point(aes(color = carat)) +
  facet_wrap(~ cut, scales = "free")
```

Try facetting by carat and grouping by cut
```{r e.7.2.7.1_plot2}
diamonds2 <- diamonds
diamonds2$carat_i <- cut_interval(diamonds$carat, 10)
  

ggplot(diamonds, aes(cut, price)) +
  geom_boxplot(aes(color = cut)) +
  coord_flip() +
  facet_wrap(~ diamonds2$carat_i)
```
Facetting cut and group by carat is better from my point of view.

**_Question 2_** Diamonds: compare the relationship between price and carat for each
colour. What makes it hard to compare the groups? Is grouping better
or facetting? If you use facetting, what annotation might you add to make
it easier to see the differences between panels?

With facetting, each group of cut is separated in its own plot, therefore, leaving no
chance for us to see the differences when every cut is on the same plot. Depends on your
personal taste, you can choose either faceting or grouping.

**_Question 3_** Why is facet wrap() generally more useful than facet grid()?

Reading from the documentation:
facet_wrap wraps a 1d sequence of panels into 2d. This is generally a better use of screen space than facet_grid because most displays are roughly rectangular.

**_Question 4_** Recreate the following plot. It facets mpg2 by class, overlaying a smooth
curve fit to the full dataset.

```{r mpg2}
mpg2 <- subset(mpg, cyl != 5 & drv %in% c("4", "f") & class != "2seater")
head(mpg2)
```
```{r}
ggplot(mpg2, aes(displ, hwy)) + 
  geom_smooth(data = select(mpg2, -class), se = FALSE) + 
  geom_point() + 
  facet_wrap(~class, nrow = 2)
```

In documentation:
The data to be displayed in this layer. There are three options:
If NULL, the default, the data is inherited from the plot data as specified in the call to ggplot.
A data.frame, or other object, will override the plot data. All objects will be fortified to produce a data frame. See fortify for which variables will be created.
A function will be called with a single argument, the plot data. The return value must be a data.frame., and will be used as the layer data.
