---
title: "ggplot2_solutions_chapter11"
author: "Nade Kang"
date: "July 14, 2018"
output: html_document
---

# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 11 Modeling for Visualization

## *Exercise 11.2.1*
### Load Packages
```{r setup, results='hide'}
library(tidyverse)
library(ggthemes)
```

**_Question 1_** What happens if you repeat the above analysis with all diamonds? (Not
just all diamonds with two or fewer carats). What does the strange geometry
of log(carat) vs relative price represent? What does the diagonal line
without any points represent?

Applying the analysis to all diamonds, without filtering diamonds with <= 2 carat.
```{r e.11.2.1.1}
diamonds_all <- diamonds %>% 
  mutate(lprice = log2(price),
         lcarat = log2(carat))
head(select(diamonds_all, carat, price, lcarat, lprice))

ggplot(diamonds_all, aes(lcarat, lprice)) +
  geom_bin2d() +
  geom_smooth(method = "lm", color = "linen", size = 2, se = FALSE) +
  ggtitle("Log Price Plotted Against Log Carat\n for All Diamonds") +
  theme(plot.title = element_text(hjust = 0.5))

mod_all <- lm(lprice ~ lcarat, data = diamonds_all)
coef(summary(mod_all))

diamonds_all <- diamonds_all %>% 
  mutate(rel_price = resid(mod_all))

ggplot(diamonds_all, aes(lcarat, rel_price)) +
  geom_bin2d() +
  ggtitle("Relative Price Plotted Against Log Carat\n for All Diamonds") +
  theme(plot.title = element_text(hjust = 0.5))

```

As the log carat increases, the relative price drops gradually below zero. This means that 
diamonds with large carat are cheapter than expected.

**_Question 2_** made an unsupported assertion that lower-quality diamonds tend to be
larger. Support my claim with a plot.

```{r e.11.2.1.2}
ggplot(diamonds_all, aes(color, carat)) +
  geom_boxplot(aes(color = cut))
```

Indeed, from this boxplot of carat against color with cut, we can clearly see that,
judging from color, D(the best) has smaller carat, while H, I and J have a larger carat.
In addition, Fair cut is generally larger than other cuts in terms of carat.

**_Question 3_** Can you create a plot that simultaneously shows the effect of colour, cut,
and clarity on relative price? If there’s too much information to show on
one plot, think about how you might create a sequence of plots to convey
the same message.

We can use facet_wrap() to help with the plot:

```{r e.11.2.1.3}
ggplot(diamonds_all, aes(cut, rel_price)) +
  geom_boxplot(aes(color = color)) +
  facet_wrap(~clarity)
```


**_Question 4_** How do depth and table relate to the relative price?

```{r e.11.2.1.4_plot1}
ggplot(diamonds_all, aes(depth, rel_price)) +
  geom_point()
```

It seems that the depth doesn't show a clear pattern wIth relative_price, although, given
a certain depth of diamond, the relative price could be high or low. This because the price
of a diamonds usually depends on color, cut, and clarity rather than on depth.

```{r e.11.2.1.4_plot2}
ggplot(diamonds_all, aes(table, rel_price)) +
  geom_point()
```

Interesting pattern for table and relative price. As we can see from the plot, with the table
increasing fom 50 to about 59, the relative price variated greater, and then smaller as table
continue to increase.

## *Exercises 11.3.1*

**_Question 1_** The final plot shows a lot of short-term noise in the overall trend. How
could you smooth this further to focus on long-term changes?
```{r}
deseas <- function(x, month) {
resid(lm(x ~ factor(month), na.action = na.exclude))
}

txhousing <- txhousing %>% 
  group_by(city) %>% 
  mutate(rel_sales = deseas(log(sales), month))

ggplot(txhousing, aes(date, rel_sales)) + 
  geom_line(aes(group = city), alpha = 1/5) + 
  geom_line(stat = "summary", fun.y = "mean", colour = "red")

ggplot(txhousing, aes(date, rel_sales)) + 
  geom_line(aes(group = city), alpha = 1/5) + 
  geom_line(stat = "summary", fun.y = "mean", colour = "red") +
  geom_smooth(method = "loess", se = FALSE)

```


**_Question 2_** If you look closely (e.g. + xlim(2008, 2012)) at the long-term trend you’ll
notice a weird pattern in 2009–2011. It looks like there was a big dip in
2010. Is this dip “real”? (i.e. can you spot it in the original data)

**_Question 3_** What other variables in the TX housing data show strong seasonal effects?
Does this technique help to remove them?

**_Question 4_** Not all the cities in this data set have complete time series. Use your dplyr
skills to figure out how much data each city is missing. Display the results
with a visualisation.

**_Question 5_** Replicate the computation that stat summary() did with dplyr so you can
plot the data “by hand”.