---
title: "ggplot2_solutions_chapter10"
author: "Nade Kang"
date: "July 13, 2018"
output: html_document
---

# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 10 Data Transformation

## *Exercise 10.2.3*
### Load Packages
```{r setup, results='hide'}
library(tidyverse)
library(ggthemes)
```

**_Question 1_** Practice your filtering skills by:
- Finding all the diamonds with equal x and y dimensions.
```{r e.10.2.3.1_1}
diamonds_xey <- filter(diamonds, x == y)
head(diamonds_xey)
```

- A depth between 55 and 70.
Here, depth refers to the value z.
Assuming the between means inclusive:
```{r e.10.2.3.1_2}
diamonds_z5570 <- diamonds %>% filter(z >= 55 & z <= 70)
head(diamonds_z5570)
```

So there is no such diamond that has a depth between 55 and 70.

- A carat smaller than the median carat.
```{r e.10.2.3.1_3}
dia_cara_lessmedian <- filter(diamonds, carat < median(diamonds$carat))
head(dia_cara_lessmedian)
```

- Cost more than $10,000 per carat
```{r e.10.2.3.1_4}
more10000pc <- filter(diamonds, price/carat > 10000)
head(more10000pc)
```


- Are of good or better quality
Are of good or better means excluding fiar:
```{r e.10.2.3.1_5}
besseralsgut <- filter(diamonds, cut != "Fair")
head(besseralsgut)
```


**_Question 2_** Fill in the question marks in this table:

Expression   TRUE   FALSE   NA
x            x
!x                  x
is.na(x)                   x
!is.na(x)    x        
is.na(x)|x   x
is.na(x)|!x  x

Didn't understand this table, waiting for help from Stack Overflow.

**_Question 3_** Repeat the analysis of outlying values with z. Compared to x and y, how
would you characterise the relationship of x and z, or y and z?

We make a first plot to take a look at the data:
```{r}
ggplot(diamonds, aes(x, z)) +
  geom_jitter()
```

Then we know to extract x > 0 and z > 0, with z < 10:
```{r e.10.2.3.3_plot1}
diamonds_zok <- filter(diamonds, x > 0, z > 0, z < 10)
ggplot(diamonds_zok, aes(x, z)) +
  geom_bin2d() +
  geom_abline(slope = 0.6, color = "white", size = 1, alpha = 0.5)
```

Despite the outliers, there seems to be a positive replationship between x and z.\

Now let's take a look at y and z:
```{r}
ggplot(diamonds, aes(y, z)) +
  geom_jitter()
```

We need to extract 0 < z < 10 and 0 < y < 15:
```{r e.10.2.3.3_plot2}
diamonds_yz <- filter(diamonds, z > 0 & z < 10, y > 0 & y < 15)
ggplot(diamonds_yz, aes(y, z)) +
  geom_bin2d() +
  geom_abline(slope = 0.65, color = "white", size = 1, alpha = 0.5)
```


**_Question 4_** Install the ggplot2movies package and look at the movies that have a
missing budget. How are they different from the movies with a budget?
(Hint: try a frequency polygon plus colour = is.na(budget).)

```{r}
library(ggplot2movies)
movies <- ggplot2movies::movies
movies
movies2 <- movies %>% 
  mutate(hasbudget = ifelse(budget == 0, "Missing Budget", "Has Budget"))
movies2
ggplot(movies2, aes(budget)) +
  geom_freqpoly(aes(color = hasbudget))
```

Not quite sure what conclusion can be drawn from here. 

**_Question 5_** What is NA & FALSE and NA | TRUE? Why? Why doesn’t NA * 0 equal zero?
What number times zero does not equal 0? What do you expect NA ˆ 0 to
equal? Why?
```{r}
NA & FALSE
# FALSE
NA | TRUE
# TRUE
```
This is because the logical expressions follow the Kleene and Priest logics, view it here:
https://en.wikipedia.org/wiki/Three-valued_logic

```{r}
NA * 0
# NA
NA ^ 0
# 1
```

The reason that NA * 0 is not equal to 0 is that x * 0 = NaN is undefined when x = Inf or x = -Inf.
But anything with power to 0 is 1.

## *Exercises 10.3.2*
**_Question 1_** Practice your variable creation skills by creating the following new variables:
- The approximate volume of the diamond (using x, y, and z).

Assuming the volume of a diamond is approximately 2/3 of a cubic with approximately the same parameters:
```{r e.10.3.2.1_1}
diamonds_q1 <- diamonds %>% 
  mutate(vol = x * y * z *(2/3))
head(select(diamonds_q1, carat, cut, x:z, vol))
```


- The approximate density of the diamond.
One carat is about 0.2 gram, and to calculate density, density is gram/cm^3:
```{r e.10.3.2.1_2}
diamonds_q1 <- diamonds_q1 %>% 
  mutate(density = carat * 0.2 / vol)
head(select(diamonds_q1, carat, cut, vol, density))
```


- The price per carat.
Price per carat is simply price/carat
```{r e.10.3.2.1_3}
diamonds_q1 <- diamonds_q1 %>% 
  mutate(pricepc = price/carat)
head(select(diamonds_q1, carat, price, vol, density, pricepc))
```


- Log transformation of carat and price.
```{r e.10.3.2.1_4}
diamonds_q1 <- diamonds_q1 %>% 
  mutate(caratl = log2(carat),
         pricel = log2(price))
head(select(diamonds_q1, carat, price, caratl, pricel))
```


**_Question 2_** How can you improve the data density of ggplot(diamonds, aes(x, z)) +
stat bin2d(). What transformation makes it easier to extract outliers?

```{r}
ggplot(diamonds, aes(x, z)) +
  stat_bin2d()
```

I am not quite sure about what the question is asking for.

**_Question 3_** The depth variable is just the width of the diamond (average of x and y)
divided by its height (z) multiplied by 100 and round to the nearest integer.
Compute the depth yourself and compare it to the existing depth variable.
Summarise your findings with a plot.

Depth is calculated as height(z) divided by width (avg of x and y) times 100. There is expression
mistake in the question.
```{r e.10.3.2.3}
diamonds_q1 <- diamonds_q1 %>% 
  mutate(depth_cal = round(z / ((x + y)/2) * 100))
head(select(diamonds_q1, depth_cal, depth))
```

Make a plot:
We calculate the difference between the given depth and our calculated depth_cal, then we make 
a plot, if the plots shows more positive values, then it means our calculation is below the original,
otherwise, above.
```{r e.10.3.2.3_plot}
diamonds_q1 <- diamonds_q1 %>% 
  mutate(diff = depth - depth_cal,
         seq = seq_along(carat))
diamonds_q1
ggplot(diamonds_q1) +
  geom_area(aes(x = seq, y = diff)) +
  xlim(800, 1500) + ylim(-2, 2)
```

You can adjust the xlim and ylim to zoom in different areas of the plot to take a look. Here only
offers one set of values for an example.

**_Question 4_** Compare the distribution of symmetry for diamonds with x > y vs. x < y.
```{r e.10.3.2.4_plot1}
diamonds_q1 %>%
  filter(x > y) %>%
  ggplot(aes(x)) +
  geom_bar(binwidth = 0.2)
```

```{r e.10.3.2.4_plot2}
diamonds_q1 %>%
  filter(y > x) %>%
  ggplot(aes(x)) +
  geom_bar(binwidth = 0.2)
```


## *Exercises 10.4.3*
**_Question 1_** For each year in the ggplot2movies::movies data determine the percent of
movies with missing budgets. Visualise the result.
```{r view, eval=FALSE}
movies <- ggplot2movies::movies
View(movies)
```

We need to create a boolean column called with_budget, so if the movie has a budget, we place TRUE,
else we place FALSE

```{r e.10.4.3.1}
movies1 <- movies %>% 
  mutate(with_budget = ifelse(is.na(budget), FALSE, TRUE)) %>%
  group_by(year) %>%
  summarise(
    hasBudget = sum(with_budget),
    total = n(),
    propNoBudget = 1 - hasBudget/total
  )
ggplot(movies1, aes(year, propNoBudget)) +
  geom_area(aes(fill = "linen")) +
  labs(fill = expression(atop("No Budget"))) +
  theme(legend.title = element_text(vjust = 0))
  
movies1
```

As you can see from the plot, the percentage of movies with missing budget decreased with
the progression of time.

**_Question 2_** How does the average length of a movie change over time? Display your
answer with a plot, including some display of uncertainty.

We only need to group the data by year, and then calculate the mean of length, then use a 
line plot (since we have time and a continuous variable).

```{r e.10.4.3.2}
q2 <- movies %>% 
  group_by(year) %>% 
  summarise(mean = mean(length))
q2
ggplot(q2, aes(year, mean)) +
  geom_line()
```

As we can see, the average length of movies increased from 1890s to now.

**_Question 3_** For each combination of diamond quality (e.g. cut, colour and clarity),
count the number of diamonds, the average price and the average size.
Visualise the results.

Assume size means depth
```{r e.10.4.3.3_cut}
diamonds_q3 <- diamonds %>% 
  group_by(cut) %>%
  summarise(
    number = n(),
    avgprice = mean(price),
    avgsize = mean(depth)
  )
diamonds_q3

ggplot(diamonds, aes(cut)) +
  geom_bar() +
  ggtitle("Number of diamonds in each cut")

ggplot(diamonds_q3, aes(cut, avgprice)) +
  geom_bar(stat = "identity") +
  ggtitle("Average Price of Each Cut")

ggplot(diamonds_q3, aes(cut, avgsize)) +
  geom_bar(stat = "identity") +
  ggtitle("Average Size of Each Cut")

```

```{r e.10.4.3.3_color}
diamonds_q3 <- diamonds %>% 
  group_by(color) %>%
  summarise(
    number = n(),
    avgprice = mean(price),
    avgsize = mean(depth)
  )

ggplot(diamonds, aes(color)) +
  geom_bar() +
  ggtitle("Number of diamonds in Each Color")

ggplot(diamonds_q3, aes(color, avgprice)) +
  geom_bar(stat = "identity") +
  ggtitle("Average Price of Each Color")

ggplot(diamonds_q3, aes(color, avgsize)) +
  geom_bar(stat = "identity") +
  ggtitle("Average Size of Each Color")
```

```{r e.10.4.3.3_clarity}
diamonds_q3 <- diamonds %>% 
  group_by(clarity) %>%
  summarise(
    number = n(),
    avgprice = mean(price),
    avgsize = mean(depth)
  )

ggplot(diamonds, aes(clarity)) +
  geom_bar() +
  ggtitle("Number of diamonds in Each Clarity")

ggplot(diamonds_q3, aes(clarity, avgprice)) +
  geom_bar(stat = "identity") +
  ggtitle("Average Price of Each Clarity")

ggplot(diamonds_q3, aes(clarity, avgsize)) +
  geom_bar(stat = "identity") +
  ggtitle("Average Size of Each Clarity")
```


**_Question 4_** Compute a histogram of carat by “hand” using a binwidth of 0.1. Display
the results with geom bar(stat = "identity"). (Hint: you might need to
create a new variable first.)

Not sure about the best way to compute by hand.

```{r e.10.4.3.4}
ggplot(diamonds, aes(carat)) +
  geom_histogram(binwidth = 0.1)
```


**_Question 5_** In the baseball example, the batting average seems to increase as the
number of at bats increases. Why?

Not familiar with baseball, but practice makes perfect?

## *Exercises 10.5.1*
**_Question 1_** Translate each of the examples in this chapter to use the pipe.

Most of my codes above used pipes, but each example is too much.

**_Question 2_** What does the following pipe do?
```{r, eval=FALSE}
library(magrittr)
x <- runif(100)
x %>%
subtract(mean(.)) %>%
raise_to_power(2) %>%
mean() %>%
sqrt()
```

It loads the magittr package first, then assign 100 uniformly distributed random variable between
0 and 1 to x. Using this x, subtract each value by the mean of x. Then raise the value to the power
of 2, then calculate the mean of the result, then square the result. As you can see, this is calculating
the VARIANCE.

**_Question 3_** Which player in the Batting dataset has had the most consistently good
performance over the course of their career?

Not familiar with baseball, and choose the criteria, group the data by player, arrange criteria in decending order.