---
title: "ggplot2_solutions_chapter11"
author: "Nade Kang"
date: "July 14, 2018"
output: html_document
---

# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 11 Modeling for Visualization

## *Exercise 11.2.1*
### Load Packages
```{r setup, results='hide'}
library(tidyverse)
library(ggthemes)
```

**_Question 1_** What happens if you repeat the above analysis with all diamonds? (Not
just all diamonds with two or fewer carats). What does the strange geometry
of log(carat) vs relative price represent? What does the diagonal line
without any points represent?

Applying the analysis to all diamonds, without filtering diamonds with <= 2 carat.
```{r e.11.2.1}
diamonds_all <- diamonds %>% 
  mutate(lprice = log2(price),
         lcarat = log2(carat))
head(select(diamonds_all, carat, price, lcarat, lprice))

ggplot(diamonds_all, aes(lcarat, lprice)) +
  geom_bin2d() +
  geom_smooth(method = "lm", color = "linen", size = 2, se = FALSE) +
  ggtitle("Log Price Plotted Against Log Carat\n for All Diamonds") +
  theme(plot.title = element_text(hjust = 0.5))

mod_all <- lm(lprice ~ lcarat, data = diamonds_all)
coef(summary(mod_all))

diamonds_all <- diamonds_all %>% 
  mutate(rel_price = resid(mod_all))

ggplot(diamonds_all, aes(lcarat, rel_price)) +
  geom_bin2d() +
  ggtitle("Relative Price Plotted Against Log Carat\n for All Diamonds") +
  theme(plot.title = element_text(hjust = 0.5))

```

As the log carat increases, the relative price drops gradually below zero. This means that 
diamonds with large carat are cheapter than expected.

**_Question 2_** made an unsupported assertion that lower-quality diamonds tend to be
larger. Support my claim with a plot.

```{r e.11.2.2}
ggplot(diamonds_all, aes(color, carat)) +
  geom_boxplot(aes(color = cut))
```

Indeed, from this boxplot of carat against color with cut, we can clearly see that,
judging from color, D(the best) has smaller carat, while H, I and J have a larger carat.
In addition, Fair cut is generally larger than other cuts in terms of carat.

**_Question 3_** Can you create a plot that simultaneously shows the effect of colour, cut,
and clarity on relative price? If there’s too much information to show on
one plot, think about how you might create a sequence of plots to convey
the same message.

We can use facet_wrap() to help with the plot:

```{r e.11.2.2}
ggplot(diamonds_all, aes(cut, rel_price)) +
  geom_boxplot(aes(color = color)) +
  facet_wrap(~clarity)
```


**_Question 4_** How do depth and table relate to the relative price?