---
title: "ggplot2_solutions_chapter10"
author: "Nade Kang"
date: "July 13, 2018"
output: html_document
---

# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 10 Data Transformation

## *Exercise 10.2.3*
### Load Packages
```{r setup, results='hide'}
library(tidyverse)
library(ggthemes)
```

**_Question 1_** Practice your filtering skills by:
- Finding all the diamonds with equal x and y dimensions.
```{r e.10.2.3.1_1}
diamonds_xey <- filter(diamonds, x == y)
head(diamonds_xey)
```

- A depth between 55 and 70.
Here, depth refers to the value z.
Assuming the between means inclusive:
```{r e.10.2.3.1_2}
diamonds_z5570 <- diamonds %>% filter(z >= 55 & z <= 70)
head(diamonds_z5570)
```

So there is no such diamond that has a depth between 55 and 70.

- A carat smaller than the median carat.
```{r e.10.2.3.1_3}
dia_cara_lessmedian <- filter(diamonds, carat < median(diamonds$carat))
head(dia_cara_lessmedian)
```

- Cost more than $10,000 per carat
```{r e.10.2.3.1_4}
more10000pc <- filter(diamonds, price/carat > 10000)
head(more10000pc)
```


- Are of good or better quality
Are of good or better means excluding fiar:
```{r e.10.2.3.1_5}
besseralsgut <- filter(diamonds, cut != "Fair")
head(besseralsgut)
```


**_Question 2_** Fill in the question marks in this table:

Expression   TRUE   FALSE   NA
x            x
!x                  x
is.na(x)                   x
!is.na(x)           x

Didn't understand this table, waiting for help from Stack Overflow.

**_Question 3_** Repeat the analysis of outlying values with z. Compared to x and y, how
would you characterise the relationship of x and z, or y and z?

We make a first plot to take a look at the data:
```{r}
ggplot(diamonds, aes(x, z)) +
  geom_jitter()
```

Then we know to extract x > 0 and z > 0, with z < 10:
```{r e.10.2.3.3_plot1}
diamonds_zok <- filter(diamonds, x > 0, z > 0, z < 10)
ggplot(diamonds_zok, aes(x, z)) +
  geom_bin2d() +
  geom_abline(slope = 0.6, color = "white", size = 1, alpha = 0.5)
```

Despite the outliers, there seems to be a positive replationship between x and z.\

Now let's take a look at y and z:
```{r}
ggplot(diamonds, aes(y, z)) +
  geom_jitter()
```

We need to extract 0 < z < 10 and 0 < y < 15:
```{r e.10.2.3.3_plot2}
diamonds_yz <- filter(diamonds, z > 0 & z < 10, y > 0 & y < 15)
ggplot(diamonds_yz, aes(y, z)) +
  geom_bin2d() +
  geom_abline(slope = 0.65, color = "white", size = 1, alpha = 0.5)
```


**_Question 4_** Install the ggplot2movies package and look at the movies that have a
missing budget. How are they different from the movies with a budget?
(Hint: try a frequency polygon plus colour = is.na(budget).)

```{r}
library(ggplot2movies)
movies <- ggplot2movies::movies
movies
movies2 <- movies %>% 
  mutate(hasbudget = ifelse(budget == 0, "Missing Budget", "Has Budget"))
movies2
ggplot(movies2, aes(budget)) +
  geom_freqpoly(aes(color = hasbudget))
```

Not quite sure what conclusion can be drawn from here. 

**_Question 5_** What is NA & FALSE and NA | TRUE? Why? Why doesn’t NA * 0 equal zero?
What number times zero does not equal 0? What do you expect NA ˆ 0 to
equal? Why?
```{r}
NA & FALSE
# FALSE
NA | TRUE
# TRUE
```
This is because the logical expressions follow the Kleene and Priest logics, view it here:
https://en.wikipedia.org/wiki/Three-valued_logic

```{r}
NA * 0
# NA
NA ^ 0
# 1
```

The reason that NA * 0 is not equal to 0 is that x * 0 = NaN is undefined when x = Inf or x = -Inf.
But anything with power to 0 is 1.