---
title: "ggplot2_solutions_chapter5"
author: "Nade Kang"
date: "July 9, 2018"
output: html_document
---

# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 5 Build a Plot Layer by Layer

## *Exercise 5.3.1*
### Load Packages
```{r setup, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

**_Question 1_** The first two arguments to ggplot are data and mapping. The first two
arguments to all layer functions are mapping and data. Why does the order
of the arguments differ? (Hint: think about what you set most commonly.)

ggplot() sets the data first, while geom layers set their aesthetics first.

**_Question 2_** The following code uses dplyr to generate some summary statistics about
each class of car (you’ll learn how it works in Chap. 10).

```{r}
library(dplyr)
class <- mpg %>% 
  group_by(class) %>% 
  summarise(n = n(), hwy = mean(hwy))
class
```

```{r e.5.3.1.2}
ggplot(mpg, aes(class, hwy)) +
  geom_jitter(width = 0.05, size = 2) +
  geom_point(aes(y = hwy), data = class, size = 4, color = "red") +
  geom_text(aes(y = 10, label = paste0("n = ", n)), data = class)
```

The process behind the scene:
First, you use the ggplot() function to define the main dataset and aesthetics you want to use. Here, we want to plot the hwy(y-axis) as points against class(x-axis). This is the first line of code.

Second, remember, ggplot2 works by plotting a graph layer by layer. The second line of code is the first layer that we are going to add. Instead of a scatterplot using geom_point(), we want to use jitter plot to avoid overplotting in geom_point(). Here, we are using the same dataset - mpg, and using the same aesthetics (x = class, y = hwy) as we indicated in ggplot(), so we do not change anything. But, to make it look similar to the original plot in the textbook, we want to squeeze the width = 0.05 (or you can adjust this number to 0.1, 0.2 to take a look), and we can add a size = 2 or 1 or 3, it doesn't matter.

Third, we want to create the same red dots in the original graph. This red dot, is the third layer. Here I used geom_point() because these red dots are basically points. However, red dots means the mean value of the y values on corresponding x values. In the original dataset mpg, we do not have this variable. So, here in the second layer of red dots, we need to reset some of the aesthetics and data. We want aesthetics to have y=hwy, this is the same as the ggplot(), but we change the dataset so data = class. Now, the layer knows that each class matches 1 y-axis value which is the hwy as mean value. We have the red dots there, but we also want to shape the size so it looks big, and color to red.

Finally, to add the layer of labels. Any annotation, we use geom_text() layer to do it. We set the aes(y = 10), because this sets the height of the label near position of y = 10. Then we want to set the label value. But here, we only know the n, which is the count. But in the original graph, it is n = integer. So we need to use the paste0() function to concatenate string and numbers. write, label = paste0("n = ", n) to have an effect of n = integer, then, we set the data = class, because we are not using the original mpg.

## *Exercise 5.4.3*
**_Question 1_** Simplify the following plot specifications:
There is a type in the textbook, the variable is displ, not disp.
```{r, eval=FALSE}
ggplot(mpg) + geom_point(aes(mpg$displ, mpg$hwy)) 

ggplot() + geom_point(mapping = aes(y = hwy, x = cty), data = mpg) + geom_smooth(data = mpg, mapping = aes(cty, hwy))

ggplot(diamonds, aes(carat, price)) + geom_point(aes(log(brainwt), log(bodywt)), data = msleep)
```

```{r e.5.4.3.1_1}
ggplot(mpg, aes(displ, hwy)) + geom_point()
```

```{r e.5.4.3.1_2}
ggplot(mpg) + geom_point(aes(cty, hwy)) + geom_smooth(aes(cty, hwy))
```

```{r e.5.4.3.1_3}
ggplot(aes(log(brainwt), log(bodywt)), data = msleep) + geom_point()
```


**_Question 2_** What does the following code do? Does it work? Does it make sense?
Why/why not?

```{r}
ggplot(mpg) + geom_point(aes(class, cty)) + geom_boxplot(aes(trans, hwy)) + coord_flip()
```

This plot is using mpg dataset as a blank background without setting aesthetics. Then
it adds the first layer using scatterplot with class as x-axis and cty on y axis.
After that, it adds another layer using boxplot with trans on x-axis and hwy on y-axis.

It doesn't work because the x-axis: 1. too crowded, need to add coord_flip() to read all
the class and trans labels. 2. Trans and class are mixed together, it doesn't make sense.
Also, boxplot is mixed with scatterplot. Y-axis has two different value groups, the hwy 
and the cty. These two contain both continuous values but are different in nature because one
is city miles per gallon, the other is highway miles per gallon.
In the end, the x-axis and y-axis labels only have class and cty, it doesn't label trans
and hwy.

**_Question 3_** What happens if you try to use a continuous variable on the x axis in one
layer, and a categorical variable in another layer? What happens if you do
it in the opposite order?

```{r e.5.4.3_1, eval=FALSE}
ggplot(mpg) +
  geom_point(aes(hwy, cyl)) +
  geom_point(aes(drv, cty))
```

```{r e.5.4.3_2}
ggplot(mpg) +
  geom_point(aes(drv, cty)) +
  geom_point(aes(hwy, cyl))
```

If you set first x-axis with categorical values, and then with continuous values, the plot
will run but the result doesn't make sense at all. However, if you do it the opposite way,
R will report errors.

## *Exercises 5.5.1*
**_Question 1_** Download and print out the ggplot2 cheatsheet from http://www.rstudio.
com/resources/cheatsheets/ so you have a handy visual reference for all
the geoms.

Easy to do, just download.

**_Question 2_** Look at the documentation for the graphical primitive geoms. Which aesthetics
do they use? How can you summarise them in a compact form?

geom_primitives such as geom_point(), geom_path(), geom_rect(), geom_polyon(), geom_text() and 
so on all have common aesthetics such as: x, y, color, alpha, fill, group, size.
Other features such as geom_point()'s stroke, or geom_path() and geom_rect()'s linetype, belong 
to these funtions themselves. But the common ones are available in each function.

**_Question 3_** What’s the best way to master an unfamiliar geom? List three resources
to help you get started.

You can use Winston Chang's R Graphics Cookbook, which shows you the recipe for different types of plots.
Or you could use Hadley Wickham's gplot2 Elegant Graphics for Data Analysis.
Or you can just use ?function to read the R documentation with examples to learn.

**_Question 4_** For each of the plots below, identify the geom used to draw it.
The first one on the upper left is violin
```{r e.5.5.1.4_plot1}
ggplot(mpg, aes(drv, displ)) +
  geom_violin(size = 1, color = "black")
```

The one on upper right is a jitter plot with size and alpha features. It is not geom_point()
because this results in overplotting.
```{r e.5.5.1.4_plot2}
ggplot(mpg) +
  geom_jitter(aes(x = hwy, y = cty, alpha = 0.1, size = cyl)) +
  theme(legend.position = "none")
```

The middle left plot is a hexagonal plot
```{r e.5.5.1.4_plot3}
ggplot(mpg) +
  geom_hex(aes(x = hwy, y = cty))
```

Middle right is a jitter plot, thanks to stack overflow user PoGibas for the help. PoGibas mentioned:
"By default jitter in geom_jitter is too large and we need to specify our own height and width of jitter by using position_jitter function.""
```{r e.5.5.1.4_plot4}
ggplot(mpg, aes(x = cyl, y = drv)) +
  geom_jitter(width = 0.05, height = 0.05)
# or geom_jitter(position = position_jitter(0.05, 0.05))
```

Next, we have to use the economics package in ggplot2.
```{r e.5.5.1.4_plot5}
ggplot(economics, aes(date, psavert)) +
  geom_area()
```

Finally, the last plot.
```{r e.5.5.1.4_plot6}
ggplot(economics, aes(uempmed, psavert)) +
  geom_path(size = 1)
```

**_Question 5_** For each of the following problems, suggest a useful geom:
-Display how a variable has changed over time.
geom_line()
-Show the detailed distribution of a single variable.
geom_histogram()
-Focus attention on the overall trend in a large dataset.
geom_line(), geom_area()
-Draw a map.
geom_sf(), geom_polygon(), coord_quickmap()    
-Label outlying points.
geom_point(), geom_text()

## *Exercises 5.6.2*
**_Question 1_** The code below creates a similar dataset to stat smooth(). Use the appropriate
geoms to mimic the default geom smooth() display.

```{r}
mod <- loess(hwy ~ displ, data = mpg)
smoothed <- data.frame(displ = seq(1.6, 7, length = 50))
pred <- predict(mod, newdata = smoothed, se = TRUE)
smoothed$hwy <- pred$fit
smoothed$hwy_lwr <- pred$fit - 1.96 * pred$se.fit
smoothed$hwy_upr <- pred$fit + 1.96 * pred$se.fit
smoothed
```

We need to use geom_ribbon() instead of geom_area(), though area sounds more like coloring the region between
the two standard deviation values. geom_ribbon() and geom_area() both display a y interval for given x values.
However, geom_are() is a special case of geom_ribbon() for geom_area()'s ymin is fixed at 0.

The sequence of the layers, without knowing other functions to deal with it, is important. If we plot the geom_ribbon()
in the end filled with color, it will shadow the smooth line and data points.
```{r e.5.6.2.1}
ggplot(mpg, aes(displ, hwy)) +
  geom_ribbon(data = smoothed, aes(ymin = hwy_lwr, ymax = hwy_upr, color = "red"), fill = "grey") +
  geom_point(position = "jitter") +
  geom_line(data = smoothed, aes(y = hwy), size = 1, color = "red")
```


**_Question 2_** What stats were used to create the following plots?
Not quite sure about this one.

**_Question 3_** Read the help for stat sum() then use geom count() to create a plot that
shows the proportion of cars that have each combination of drv and trans.

```{r}
mpg %>% 
  group_by(drv, trans) %>% 
  summarise(n = n()) %>% 
  mutate(drvtrans = paste0(drv, "-", trans)) %>% 
  ggplot() +
  geom_bar(aes(drvtrans, n), stat = "identity") +
  coord_flip()
```

## *Exercises 5.7.1*
**_Question 1_** When might you use position nudge()? Read the documentation.

position_nudge is generally useful for adjusting the position of items on discrete scales by a small amount. Nudging is built in to geom_text because it's so useful for moving labels a small distance from what they're labelling.

e.g.:
```{r, eval=FALSE, results='hide'}
ggplot(df, aes(x, y)) +
  geom_point() +
  geom_text(aes(label = y), position = position_nudge(y = -0.1))
```

**_Question 2_** Many position adjustments can only be used with a few geoms. For example,
you can’t stack boxplots or errors bars. Why not? What properties
must a geom possess in order to be stackable? What properties must it
possess to be dodgeable?

Not sure.

**_Question 3_** Why might you use geom jitter() instead of geom count()? What are the
advantages and disadvantages of each technique?

Count() counts the number of obervations at each given x values, and then map the point.
The more observation one given x value has, the larger the point.

Jitter() adds random noise and shows how the observations are distributed in the plot.

**_Question 4_** When might you use a stacked area plot? What are the advantages and
disadvantages compared to a line plot?

When you want to separate the x values based on a given category, you use stacked area plot.
Advantage is it allows you to show the difference based on the criteria, but sometimes it is
hard to figure out the real magnitude.