---
title: "ggplot2_solutions_chapter12"
author: "Nade Kang"
date: "July 14, 2018"
output: html_document
---

# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 12 Programming with ggplot2

## *Exercise 12.2.1*
### Load Packages
```{r setup, results='hide'}
library(tidyverse)
library(ggthemes)
```

**_Question 1_** Create an object that represents a pink histogram with 100 bins.

We name this object as hist100:
```{r e.12.2.1.1}
hist100 <- geom_histogram(bins  = 100, color = "pink", fill = "pink")

# Test this object with diamonds:
ggplot(diamonds, aes(price)) +
  hist100
```


**_Question 2_** Create an object that represents a fill scale with the Blues ColorBrewer
palette.

P138 in textbook, we just use the default scale_fill_brewer() so it is blue.

```{r e.12.2.1.2}
bluePalette <- scale_fill_brewer()

# Test this object with diamonds dataset:
ggplot(diamonds, aes(cut, price, fill = cut)) +
  geom_bar(stat = "summary", fun.y = "mean") +
  scale_fill_brewer()
```


**_Question 3_** Read the source code for theme grey(). What are its arguments? How does
it work?

```{r}
theme_grey
```


**_Question 4_** Create scale colour wesanderson(). It should have a parameter to pick the
palette from the wesanderson package, and create either a continuous or
discrete scale.

Package not available for R3.4.4

