---
title: "ggplot2_solutions_chapter6"
author: "Nade Kang"
date: "July 10, 2018"
output: html_document
---

# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 6 Scales Axes and Legends

## *Exercise 6.2.1*
### Load Packages
```{r setup, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

**_Question 1_** What happens if you pair a discrete variable to a continuous scale? What
happens if you pair a continuous variable to a discrete scale?

Pair a discrete variable to continuous scale:
```{r e.6.2.1.1_plot1}
ggplot(mpg, aes(displ, hwy)) +
geom_point() +
scale_x_continuous() +
scale_y_continuous()
```
```{r e.6.2.1.1_plot2}
ggplot(mpg, aes(displ, hwy)) +
geom_point() +
scale_x_discrete() +
scale_y_discrete()
```

Let's compare the two graphs. The first plot used both scale_x_continuous() and scale_y_continuous().
The second plot used scale_x_discrete() and scale_y_discrete for both axes. The difference between
the two graphs doesn't lie within the positions of the points, but the background and the units on
the two axes. If you use discrete for continuous, you won't see the hwy units on the y-axis, nor will
you see the units for displ on x-axis.

On the other hand, we can try continuous scale on discrete variables:
```{r e.6.2.1.1_plot3}
ggplot(mpg, aes(class, hwy)) +
  geom_jitter(width = 0.05, height = 0.05)
```

Now we change the scale on the previous plot:
```{r e.6.2.1.1_plot4, eval=FALSE}
ggplot(mpg, aes(class, hwy)) +
  geom_jitter(width = 0.05, height = 0.05) +
  scale_x_continuous()
# Error: Discrete value supplied to continuous scale
```

We were not allowed to do so because of the error message: Discrete value
supplied to continuous scale.

So in conclusion, we could supply discrete scale to continuous variables, but not
vice versa.

**_Question 2_** Simplify the following plot specifications to make them easier to understand.
```{r, eval=FALSE}
ggplot(mpg, aes(displ)) +
scale_y_continuous("Highway mpg") +
scale_x_continuous() +
geom_point(aes(y = hwy))
```

The codes can be simplified as below:

```{r e.6.2.1.2_plot1}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  ylab("Highway mpg")
```

```{r, eval=FALSE}
ggplot(mpg, aes(y = displ, x = class)) +
scale_y_continuous("Displacement (l)") +
scale_x_discrete("Car type") +
scale_x_discrete("Type of car") +
scale_colour_discrete() +
geom_point(aes(colour = drv)) +
scale_colour_discrete("Drive\ntrain")
```

This can be simplified down to:
```{r e.6.2.1.2_plot2}
ggplot(mpg, aes(class, displ)) +
  geom_point(aes(color = drv)) +
  labs(x = "Type of car", y = "Displacement (l)", colour = "Drive\ntrain")
```

