---
title: "ggplot2_solutions_chapter6"
author: "Nade Kang"
date: "July 10, 2018"
output: html_document
---

# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 6 Scales Axes and Legends

## *Exercise 6.2.1*
### Load Packages
```{r setup, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

**_Question 1_** What happens if you pair a discrete variable to a continuous scale? What
happens if you pair a continuous variable to a discrete scale?

Pair a discrete variable to continuous scale:
```{r e.6.2.1.1_plot1}
ggplot(mpg, aes(displ, hwy)) +
geom_point() +
scale_x_continuous() +
scale_y_continuous()
```
```{r e.6.2.1.1_plot2}
ggplot(mpg, aes(displ, hwy)) +
geom_point() +
scale_x_discrete() +
scale_y_discrete()
```

Let's compare the two graphs. The first plot used both scale_x_continuous() and scale_y_continuous().
The second plot used scale_x_discrete() and scale_y_discrete for both axes. The difference between
the two graphs doesn't lie within the positions of the points, but the background and the units on
the two axes. If you use discrete for continuous, you won't see the hwy units on the y-axis, nor will
you see the units for displ on x-axis.

On the other hand, we can try continuous scale on discrete variables:
```{r e.6.2.1.1_plot3}
ggplot(mpg, aes(class, hwy)) +
  geom_jitter(width = 0.05, height = 0.05)
```

Now we change the scale on the previous plot:
```{r e.6.2.1.1_plot4, eval=FALSE}
ggplot(mpg, aes(class, hwy)) +
  geom_jitter(width = 0.05, height = 0.05) +
  scale_x_continuous()
# Error: Discrete value supplied to continuous scale
```

We were not allowed to do so because of the error message: Discrete value
supplied to continuous scale.

So in conclusion, we could supply discrete scale to continuous variables, but not
vice versa.

**_Question 2_** Simplify the following plot specifications to make them easier to understand.
```{r, eval=FALSE}
ggplot(mpg, aes(displ)) +
scale_y_continuous("Highway mpg") +
scale_x_continuous() +
geom_point(aes(y = hwy))
```

The codes can be simplified as below:

```{r e.6.2.1.2_plot1}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  ylab("Highway mpg")
```

```{r, eval=FALSE}
ggplot(mpg, aes(y = displ, x = class)) +
scale_y_continuous("Displacement (l)") +
scale_x_discrete("Car type") +
scale_x_discrete("Type of car") +
scale_colour_discrete() +
geom_point(aes(colour = drv)) +
scale_colour_discrete("Drive\ntrain")
```

This can be simplified down to:
```{r e.6.2.1.2_plot2}
ggplot(mpg, aes(class, displ)) +
  geom_point(aes(color = drv)) +
  labs(x = "Type of car", y = "Displacement (l)", colour = "Drive\ntrain")
```

## *Exercises 6.3.3
**_Question 1_** Recreate the following graphic:
Adjust the y axis label so that the parentheses are the right size.

```{r e.6.3.3.1}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous("Displacement",
                     breaks = c(2,3,4,5,6,7),
                     labels = c("2k", "3k", "4k", "5k", "6k", "7k")) +
  scale_y_continuous(quote(Highway (Miles/Gallon)))
```

**_Question 2_** List the three different types of object you can supply to the breaks argument.
How do breaks and labels differ?

Breaks takes one of the followings as input: 
-NULL for no breaks
-waiver() for the default breaks computed by the transformation object
-A numeric vector of positions
-A function that takes the limits as input and returns breaks as output

Labels takes one of the followings as input:
-NULL for no labels
-waiver() for the default computed by the transformation object
-A character vector giving labels (must be same length as breaks)
-A function that takes the breaks as input and returns labels as output

**_Question 3_** Recreate the following plot:

```{r e.6.3.3.3}
ggplot(mpg, aes(displ, hwy, color = drv)) +
  geom_point() +
  scale_color_discrete(labels = c("4wd", "fwd", "rwd"))
```


**_Question 4_** What label function allows you to create mathematical expressions?What
label function converts 1 to 1st, 2 to 2nd, and so on?

quote() to create mathematical expressions.
Type: scales::format and up and down arrows to check:
labels = scales::ordinal_format() to convert to 1st, 2nd, 3rd and so on.


**_Question 5_** What are the three most important arguments that apply to both axes
and legends? What do they do? Compare and contrast their operation for
axes vs. legends.
Maybe labs()? Not sure what three most important means.


## *Exercises 6.4.4*
**_Question 1_** How do you make legends appear to the left of the plot?
theme(legend.position = "left")

**_Question 2_** What’s gone wrong with this plot? How could you fix it?

```{r}
ggplot(mpg, aes(displ, hwy)) +
geom_point(aes(colour = drv, shape = drv)) +
scale_colour_discrete("Drive train")
```

The plot created two legends on the right-hand side, where you could just use one.

Here, quoting from the book:"In order for legends to be merged, they must have the same name. So if you
change the name of one of the scales, you’ll need to change it for all of them."

As a result, here, the original plot sets only colour with new name, while shape doesn't have
the same new name, the result is that these two legends cannot merge. There are two ways to fix this
plot:

```{r e.6.4.4.2_plot1}
ggplot(mpg, aes(displ, hwy)) +
geom_point(aes(color = drv, shape = drv)) +
  labs(color = "Drive train", shape = "Drive train")
```

```{r e.6.4.4.2_plot2}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = drv, shape = drv)) + 
  scale_colour_discrete("Drive train") +
  scale_shape_discrete("Drive train")
```

**_Question 3_** Can you recreate the code for this plot?

Answers are above in Question 2 answer.

## *Exercises 6.5.1*
**_Question 1_** The following code creates two plots of the mpg dataset. Modify the code
so that the legend and axes match, without using facetting!
```{r, eval=FALSE}
fwd <- subset(mpg, drv == "f")
rwd <- subset(mpg, drv == "r")
ggplot(fwd, aes(displ, hwy, colour = class)) + geom_point()
ggplot(rwd, aes(displ, hwy, colour = class)) + geom_point()
```

Use exactly expand_limits() function to set both plots' legend title to have all types
of drv with the same set of colors.

We can also set the xlim and ylim so that both plots have the same axes scales. But this
is optional.

```{r, e.6.5.1.1_plot1}
fwd <- subset(mpg, drv == "f")
rwd <- subset(mpg, drv == "r")
ggplot(fwd, aes(displ, hwy, colour = class)) + 
  geom_point() +
  scale_color_discrete("Drive train") +
  xlim(0, 10) +
  ylim(0, 45) +
  expand_limits(color = c("2seater", "compact", "midsize", "minivan",
                                   "pickup", "subcompact", "suv"))
```

```{r, e.6.5.1.1_plot2}
ggplot(rwd, aes(displ, hwy, colour = class)) + 
  geom_point() +
  scale_color_discrete("Drive train")+
  xlim(0, 10) +
  ylim(0, 45) +
  expand_limits(color = c("2seater", "compact", "midsize", "minivan",
                                   "pickup", "subcompact", "suv"))
```


**_Question 2_** What does expand limits() do and how does it work? Read the source
code.

Sometimes you may want to ensure limits include a single value, for all panels or all plots. This function is a thin wrapper around geom_blank that makes it easy to add such values.

expand_limits() solves the first question.

**_Question 3_** What happens if you add two xlim() calls to the same plot? Why?

Use question 1's plot as an example, let's test what happens when we add two xlim() calls.
```{r e.6.5.1.3}
ggplot(rwd, aes(displ, hwy, colour = class)) + 
  geom_point() +
  scale_color_discrete("Drive train")+
  xlim(0, 10) +
  xlim(0, 20) +
  ylim(0, 45) +
  expand_limits(color = c("2seater", "compact", "midsize", "minivan",
                                   "pickup", "subcompact", "suv"))
```

The result is that, the second xlim(0, 20) call overrides the first xlim(0, 10) call. As
a result, the finished plot has an x-axis ranging from 0 to 20 instead of from 0 to 10.

**_Question 4_** What does scale x continuous(limits = c(NA, NA)) do?
Let's use the same plot:
```{r e.6.5.1.4_plot1}
ggplot(fwd, aes(displ, hwy, colour = class)) + 
  geom_point() +
  xlim(0, 20)
```

```{r e.6.5.1.4_plot2}
ggplot(fwd, aes(displ, hwy, colour = class)) + 
  geom_point() +
  xlim(0, 20) +
  scale_x_continuous(limits = c(NA, NA)) 
```

This overrides the existing xlim() on the plot. So there is no more limits from xlim().

## *Exercises 6.6.5*
**_Question 1_** Compare and contrast the four continuous colour scales with the four
discrete scales.

Please refer to the textbook.

**_Question 2_** Explore the distribution of the built-in colors() using the luv colours
dataset.

```{r e.6.6.5.2}
color_df <- colors()

ggplot(luv_colours, aes(u, v)) +
  geom_point(aes(color = color_df), size = 2) +
  scale_color_identity() +
  coord_equal()

```

