---
title: "ggplot2_chapter2"
author: "Nade Kang"
date: "July 6, 2018"
output: html_document
---
# Solution Manual for ggplot2 Elegant Graphics for Data Analysis by Hadley Wickham
# ggplot2 Chpater 2 Getting Started with ggplot2

## Exercise 2.2.1
### Load Packages
```{r setup}
library(tidyverse)
```
1. List five functions that you could use to get more information about the mpg dataset.
```{r e.2.2.1.1, eval=FALSE}
View(mpg)
head(mpg)
tail(mpg)
dim(mpg)
names(mpg)
```

2. How can you find out what other datasets are included with ggplot2?
```{r e.2.2.1.2, eval=TRUE}
data(package = "ggplot2")
```

3. Apart from the US, most countries use fuel consumption (fuel consumed
over fixed distance) rather than fuel economy (distance travelled with fixed
amount of fuel). How could you convert cty and hwy into the European
standard of l/100km?
```{r e.2.2.1.3, eval=TRUE}
head(mpg)
mpg_km <- mpg %>%
  mutate(cty_l100km = cty * 235.215,
         hwy_l100lm = hwy * 235.215)
mpg_km %>% select(cty:hwy_l100lm)
```
4. Which manufacturer has the most the models in this dataset?Which model
has the most variations? Does your answer change if you remove the redundant
specification of drive train (e.g. “pathfinder 4wd”, “a4 quattro”)
from the model name?

Manufacturer with the most models
```{r}
mpg %>%
  group_by(manufacturer) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```
As you can see, Dodge has the most models, a number of 37.

